NumPy---Array Computation for Python

### 2020-01-08 body.tex
Revisions to recent developments

\begin{figure}
  \centering
  \includegraphics[width=0.5\linewidth]{static/broadcasting}
  \caption{
    Broadcasting a $1 \times 3$ array against a $2 \times 1$ array
    yields a $2 \times 3$ result.
  }
  \label{fig:broadcasting}
\end{figure}

@@

For many years, NumPy had no dedicated funding, and development time
was mostly contributed freely by students and researchers in their
spare time.  In-person meetings were sponsored off of individual's
research grants, and while industry sponsored contributions were made,
notably those by Mark Wiebe when he was sponsored by Travis Oliphant
at Enthought, the package's rapid adoption in scientific
computing took place without external investment.

The project entered a new phase of development in 2018 with the hiring of two (and later four) full-time developers.  With this new resource, NumPy was able to address long-outstanding technical debt, review more volunteer work, and implement several larger features.

In an open source project that has relied on volunteer effort for so many years, the introduction of paid work can be disruptive. In such communities, decision making power is often tied to contributed effort, and when some individuals are paid it skews the balance. To guard against this, the project set up an explicit governance structure, formalized the feature proposal process, and constructed a development roadmap with extensive feedback from the community.   In addition, individuals were hired that were already active in NumPy and its surrounding communities: they already shared the project's norms and values.  Funded developers were also expected to review code as a part of their job, thereby supporting the work done by volunteers.

Still, funded development allowed for the implementation of several bigger features, which required a sizeable amount of time, conversations with multiple stakeholders, as well as longer term focus, and were therefore difficult to accomplish by volunteer effort alone.

New work divides roughly into two categories: improving the library's interaction with external packages, and, with the benefit of several years' worth of experience, redesigning core parts of the library.

@@

>>>>

\begin{figure}
  \centering
  \includegraphics[width=0.5\linewidth]{static/broadcasting}
  \caption{\fixme{PLACEHOLDER: Stefan is working on some new figures to illustrate arrays, broadcasting, etc.}
   }
  \label{fig:broadcasting}
\end{figure}

@@

%For many years, NumPy had no dedicated funding, and development time
%was mostly contributed freely by students and researchers in their
%spare time.  In-person meetings were sponsored off of individual's
%research grants, and while industry sponsored contributions were made,
%notably those by Mark Wiebe when he was sponsored by Travis Oliphant
%at Enthought, the package's rapid adoption in scientific
%computing took place without external investment.

NumPy was initially developed by students, faculty, and researchers in their
spare time to provide a modern array computation library for Python.
They were influenced by their experiences with powerful interactive programming
languages for scientific computing like Yurik \cite{munro1995using} as well
as commercial languages like IDL and Matlab.
Often they developed code to solve their own or their colleagues problems.

As Python's popularity for scientific computing, data science, and machine
learning increased, many new use cases arose.
At the same time, the massive number of scientific Python packages that
build on NumPy mean that it has an unusually high need for stability.

In 2018, we hired two (and later four) full-time developers.
This enabled us to implement several bigger features,
which required a sizeable amount of time, conversations with multiple
stakeholders, as well as longer term focus, and were therefore difficult to
accomplish by volunteer effort alone.
To ensure that this paid effort did not disrupt our community owned character
we formalized the feature proposal process, and constructed a development
roadmap with extensive input and feedback from the community.   In addition,
we hired people who were already active in NumPy and its surrounding communities:
they already shared the project's norms and values.
Funded developers also regularly reviewed other people's code as a part of
their job, thereby supporting the work done by volunteers.

Much of the recent work falls into two categories: improving the library's
interaction with external packages, and, with the benefit of several years'
worth of experience, redesigning core parts of the library.

@@

\fixme{To be written.}

### 2020-01-08 summary.tex
Update abstract
Revisions to recent developments

% - 2-3 sentences of basic-level introduction to the field;
Software libraries for array computation provide powerful notation
for accessing, manipulating, and operating on data in vectors,
matrices, and higher-dimensional arrays \cite{iverson1980notation}.
% Stéfan: why "aggregate" data?
% - a brief account of the background and rationale of the work;
NumPy is the fundamental array computation library for the Python ecosystem
\cite{dubois2007guest,oliphant2007python,millman2011python,perez2011python}.
Not surprisingly, many important scientific research projects and results
depend on NumPy.
It plays an essential role in research analysis pipelines in fields as
diverse as physics, biology, astronomy, neuroscience, material science,
engineering, and chemistry.
For example, in astronomy, NumPy was fundamental to discovering gravitational
waves\cite{pycbc}, the first imaging of a black
hole\cite{eht-imaging}, and continues underpinning the future exploration of our universe\cite{jenness2018lsst}.
It is also one of the most commonly used machine learning tools among
enterprises \cite{451report2018}.
% - a statement of the main conclusions (introduced by the phrase 'Here we show' or its equivalent);
Here we provide an overview of the project and library organization, highlight
some recent technical developments, and outline our future plans.
% - 2-3 sentences putting the main findings into general context so it is
%   clear how the results described in the paper have moved the field forwards.
NumPy continues to evolve to meet the needs of its community, to remain
relevant in an ever-changing computing landscape, and to be a fundamental
pillar of support for the next decade of data science in Python.

>>>>

% - 2-3 sentences of basic-level introduction to the field;
Array computation libraries provide powerful notation for accessing,
manipulating, and operating on data in vectors, matrices, and
higher-dimensional arrays \cite{iverson1980notation}.
% - a brief account of the background and rationale of the work;
NumPy is the fundamental array computation library for the Python ecosystem
\cite{dubois2007guest,oliphant2007python,millman2011python,perez2011python}.
It has been actively developed for over two decades and has more than 850
contributors, nearly 300,000 dependent repositories on GitHub, and millions of
downloads per year.
It plays an essential role in research analysis pipelines in fields as
diverse as physics, biology, astronomy, neuroscience, material science,
engineering, and chemistry.
For example, in astronomy, NumPy was fundamental to discovering gravitational
waves\cite{pycbc}, the first imaging of a black hole\cite{eht-imaging}, and
continues underpinning the future exploration of our
universe\cite{jenness2018lsst}.
It is one of the most commonly used machine learning tools among
enterprises \cite{451report2018}.
% - a statement of the main conclusions (introduced by the phrase 'Here we show' or its equivalent);
Here we show how a small group of students and scientists created this
popular tool, explain the core ideas that helped make it it successful,
and highlight recent development to better serve new use cases
and future improvements.
% - 2-3 sentences putting the main findings into general context so it is
%   clear how the results described in the paper have moved the field forwards.
NumPy continues to evolve to meet the needs of its community, to remain
relevant in an ever-changing computing landscape, and to be a fundamental
pillar of support for the next decade of scientific computing, data science,
and machine learning in Python.
